---
title: "Boxplot_function"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Boxplot_function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bim)
```

## Preliminaries 

Loading packages need for data wrangling. 

```{r}
library(tidyverse)
```

```{r}
# Loading in .csv files for brain and meninges cell count data
data("UT27_Meninges_Output_Data", package = "bim")
data("UT27_Brain_Output_Data", package = "bim")

meninges <- UT27_Meninges_Output_Data
brain <- UT27_Brain_Output_Data

# selecting the variables we will be working with for brain and meninges cell count data
brain_s <- brain %>%
  select(Date, `Mouse #`,`Cage #`, ZT, Sex, Group, `Microglia Counts`, `Neutrophil Counts`, `T-Cell Counts`, `CD8+ T Cell Counts`, `CD4+ T Cell Counts`, `B Cell Counts`)

meninges_s <- meninges %>%
  select(Date, `Mouse #`,`Cage #`, ZT, Sex, Group, `Neutrophil Cells`, `T Cell Cells`, `CD8 Cells`, `CD4 Cells`, `B Cell Cells`)

# renaming variables for brain and meninges cell count data : Added "_" to column names - Stephen
meninges_select <- meninges_s %>%
  rename(
    `Meninges_Neutrophil_Counts` = `Neutrophil Cells`,
    `Meninges_T_Cell_Counts` = `T Cell Cells`,
    `Meninges_CD8+_T_Cell_Counts` = `CD8 Cells`,
    `Meninges_CD4+_T_Cell_Counts` = `CD4 Cells`,
    `Meninges_B_Cell_Counts` = `B Cell Cells`
  )

brain_select <- brain_s %>%
  rename(
    `Brain_Microglia_Counts` = `Microglia Counts`,
    `Brain_Neutrophil_Counts` = `Neutrophil Counts`,
    `Brain_T_Cell_Counts` = `T-Cell Counts`,
    `Brain_CD8+_T_Cell_Counts` = `CD8+ T Cell Counts`,
    `Brain_CD4+_T_Cell_Counts` = `CD4+ T Cell Counts`,
    `Brain_B_Cell_Counts` = `B Cell Counts`
  )

# creating a new variable that contains the cage number and the mouse number 
brain_select$`Mouse #` <- paste(brain_select$`Cage #`, brain_select$`Mouse #`, sep = "-")
meninges_select$`Mouse #` <- paste(meninges_select$`Cage #`, meninges_select$`Mouse #`, sep = "-")

# combining all the data into one data file 
combined_data <- brain_select %>%
  full_join(meninges_select)

# omitting any blank data 
combined_data <- na.omit(combined_data)


# Extra data wrangling - Stephen
# Breaks down the group into separate variables (Genotype, Age, and Age Class)
# Also ZT needs to be a factor
combined_data2 <- combined_data |> 
  mutate(ZT  = as.factor(ZT), # turns ZT into a factor 
         Genotype = case_when(grepl("WT", Group, ignore.case = TRUE) ~ "WT", # new Genotype variable ; WT or 3xTg 
                              grepl("3xTg", Group, ignore.case = TRUE) ~ "3xTg"), 
         # (?<=-) look after a dash ; [0-9]+ matches the numbers after ; $ to look at the end of the string
         Age = str_extract(Group, "(?<=-)[0-9]+$"), # new Age variable ; 6, 9, or 18
         Age_Class = case_when( # new Age Class variable ; Adult, Middle Aged, and Aged
           Age == 6 ~ "Adult",
           Age == 9 ~ "Middle Aged", 
           Age == 18 ~ "Aged")) |>
  relocate(Genotype, Age,  Age_Class, .after = Group) |> # moves new variables to the front to be with the rest of the categorical variables 
  filter(Age_Class %in% c("Adult", "Aged"),  # we were only interested in adult vs aged in WT
         Genotype == "WT")

knitr::kable(combined_data2)

```



## BOXPLOT FUNCTION EXAMPLE
This function is set up to to take strings "text" or character vectors (c("text", "text"). This is helpful if spaces are used in the variables names. 

This function creates a boxplot for each of the numerical (cell) y-variables that are provided (e.g., "Brain_T_Cell_Counts", "Brain_CD8+_T_Cell_Counts", "Meninges_CD8+_T_Cell_Counts"). These (cell) y-variables are plotted against a single categorical (group) x-variable (e.g., "Age_Class"). There is an option to choose a fill variable to separate and color the boxplots (e.g., by "Genotype") and a facet variable to separate the plots into panels (e.g., by "Sex"). There are also options to add individual data points (e.g. TRUE adds points) and to adjust the whiskers (e.g., 1.5 * IQR is the default). 

```{r}
boxplot_example <- graph_boxplots(
  data = combined_data2, # data 
  x_variable = "Age_Class", # categorical (group) variable
  y_variables = c("Brain_T_Cell_Counts", "Brain_CD8+_T_Cell_Counts", "Meninges_CD8+_T_Cell_Counts"), # numerical (cell) variables
  fill_variable = "ZT",  # fill variable
  facet_variable = "Sex", # facet variable
  add_points = TRUE, # add points 
  coef = 1.5) # whisker adjustment 

boxplot_example
```
