---
title: "Choose_best_model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choose_best_model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bim)
```

## Read in dataframes and data wrangling

```{r}
library(readr)
library(dplyr)
library(skimr)

data("UT27_Meninges_Output_Data")
meninges <- UT27_Meninges_Output_Data

data("UT27_Brain_Output_Data")
brain <- UT27_Brain_Output_Data

brain_select <- brain %>%
  select(Date, `Mouse #`,`Cage #`, ZT, Sex, Group, `Microglia Counts`, `Neutrophil Counts`, `T-Cell Counts`, `CD8+ T Cell Counts`, `CD4+ T Cell Counts`, `B Cell Counts`)

meninges_select <- meninges %>%
  select(Date, `Mouse #`,`Cage #`, ZT, Sex, Group, `Neutrophil Cells`, `T Cell Cells`, `CD8 Cells`, `CD4 Cells`, `B Cell Cells`)

meninges_select <- meninges_select %>%
  rename(
    `Meninges Neutrophil Counts` = `Neutrophil Cells`,
    `Meninges T-Cell Counts` = `T Cell Cells`,
    `Meninges CD8+ T Cell Counts` = `CD8 Cells`,
    `Meninges CD4+ T Cell Counts` = `CD4 Cells`,
    `Meninges B Cell Counts` = `B Cell Cells`
  )

brain_select <- brain_select %>%
  rename(
    `Brain Microglia Counts` = `Microglia Counts`,
    `Brain Neutrophil Counts` = `Neutrophil Counts`,
    `Brain T-Cell Counts` = `T-Cell Counts`,
    `Brain CD8+ T Cell Counts` = `CD8+ T Cell Counts`,
    `Brain CD4+ T Cell Counts` = `CD4+ T Cell Counts`,
    `Brain B Cell Counts` = `B Cell Counts`
  )

brain_select$`Mouse #` <- paste(brain_select$`Cage #`, brain_select$`Mouse #`, sep = "-")
meninges_select$`Mouse #` <- paste(meninges_select$`Cage #`, meninges_select$`Mouse #`, sep = "-")

combined_data <- brain_select %>%
  full_join(meninges_select) 

combined_data <- na.omit(combined_data)
combined_data <- combined_data %>%
  filter(!(Group %in% c("2-WT-9", "3-3xTg-9", "5-3xTg-18")))
```

Coefficient Heatmap
DESCRIPTION: This will generate a heatmap that gives an overview of which cell types correlate with which cell types. Ones that positively correlate will be represented by a blue box, negatively correlate by a red box, and don't significantly correlate by a grey box. In this example, we can see that all T cell and CD8+ T cell counts correlate with each other, even between tissues. This should inform which are the best predictors to use in a model.

```{r}
library(corrplot)
numeric_data <- combined_data[, sapply(combined_data, is.numeric)]
corrplot::corrplot(cor(numeric_data, use = "complete.obs"), method = "color")
```

Ex. using this model
DESCRIPTION: This will do a linear model of one variable, using every variable listed as a predictor and output that as the "full_model". It will the run AICc and select variables that are used as predictors in every AICc model, and only use those to re-model the data, outputting that as "reduced_model"

```{r}
library(MuMIn)

response <- "`Brain CD8+ T Cell Counts`"
predictors <- c(
  "`Meninges Neutrophil Counts`", "`Meninges T-Cell Counts`",
  "`Meninges CD8+ T Cell Counts`", "`Meninges CD4+ T Cell Counts`",
  "`Meninges B Cell Counts`", "`Brain Microglia Counts`",
  "`Brain Neutrophil Counts`", "`Brain T-Cell Counts`",
  "`Brain CD4+ T Cell Counts`", "`Brain B Cell Counts`"
)

model_results <- choose_best_model(combined_data, response, predictors)

# Access outputs:
summary(model_results$full_model)
summary(model_results$reduced_model)
model_results$top_predictors

#Save outputs
brain_CD8_model <- model_results$full_model
brain_CD8_model_reduced <- model_results$reduced_model
```

