---
title: "Descriptive_stats_function"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Descriptive_stats_function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bim)
```

## Preliminaries 

Loading packages need for data wrangling. 

```{r}
library(tidyverse)
```

## Data wrangling 
```{r}
# Loading in .csv files for brain and meninges cell count data
data("UT27_Meninges_Output_Data", package = "bim")
data("UT27_Brain_Output_Data", package = "bim")

meninges <- UT27_Meninges_Output_Data
brain <- UT27_Brain_Output_Data

# selecting the variables we will be working with for brain and meninges cell count data
brain_s <- brain %>%
  select(Date, `Mouse #`,`Cage #`, ZT, Sex, Group, `Microglia Counts`, `Neutrophil Counts`, `T-Cell Counts`, `CD8+ T Cell Counts`, `CD4+ T Cell Counts`, `B Cell Counts`)

meninges_s <- meninges %>%
  select(Date, `Mouse #`,`Cage #`, ZT, Sex, Group, `Neutrophil Cells`, `T Cell Cells`, `CD8 Cells`, `CD4 Cells`, `B Cell Cells`)

# renaming variables for brain and meninges cell count data : Added "_" to column names - Stephen
meninges_select <- meninges_s %>%
  rename(
    `Meninges_Neutrophil_Counts` = `Neutrophil Cells`,
    `Meninges_T_Cell_Counts` = `T Cell Cells`,
    `Meninges_CD8+_T-Cell_Counts` = `CD8 Cells`,
    `Meninges_CD4+_T-Cell_Counts` = `CD4 Cells`,
    `Meninges_B_Cell_Counts` = `B Cell Cells`
  )

brain_select <- brain_s %>%
  rename(
    `Brain_Microglia_Counts` = `Microglia Counts`,
    `Brain_Neutrophil_Counts` = `Neutrophil Counts`,
    `Brain_T_Cell_Counts` = `T-Cell Counts`,
    `Brain_CD8+_T_Cell_Counts` = `CD8+ T Cell Counts`,
    `Brain_CD4+_T_Cell_Counts` = `CD4+ T Cell Counts`,
    `Brain_B_Cell_Counts` = `B Cell Counts`
  )

# creating a new variable that contains the cage number and the mouse number 
brain_select$`Mouse #` <- paste(brain_select$`Cage #`, brain_select$`Mouse #`, sep = "-")
meninges_select$`Mouse #` <- paste(meninges_select$`Cage #`, meninges_select$`Mouse #`, sep = "-")

# combining all the data into one data file 
combined_data <- brain_select %>%
  full_join(meninges_select)

# omitting any blank data 
combined_data <- na.omit(combined_data)


# Extra data wrangling - Stephen
# Breaks down the group into separate variables (Genotype, Age, and Age Class)
# Also ZT needs to be a factor
combined_data2 <- combined_data |> 
  mutate(ZT  = as.factor(ZT), # turns ZT into a factor 
         Genotype = case_when(grepl("WT", Group, ignore.case = TRUE) ~ "WT", # new Genotype variable ; WT or 3xTg 
                              grepl("3xTg", Group, ignore.case = TRUE) ~ "3xTg"), 
         # (?<=-) look after a dash ; [0-9]+ matches the numbers after ; $ to look at the end of the string
         Age = str_extract(Group, "(?<=-)[0-9]+$"), # new Age variable ; 6, 9, or 18
         Age_Class = case_when( # new Age Class variable ; Adult, Middle Aged, and Aged
           Age == 6 ~ "Adult",
           Age == 9 ~ "Middle Aged", 
           Age == 18 ~ "Aged")) |>
  relocate(Genotype, Age,  Age_Class, .after = Group) |> # moves new variables to the front to be with the rest of the categorical variables 
  filter(Age_Class %in% c("Adult", "Aged"),  # we were only interested in adult vs aged in WT
         Genotype == "WT")

knitr::kable(combined_data2)


```

## DESCRIPTIVE STATS FUNCTION EXAMPLE 
This function is set up to to take strings "text" or character vectors (c("text", "text"). This is helpful if spaces are used in the variables names. 

This function creates a table containing the categorical (group) variables that are provided (e.g.,"Sex","ZT", "Age_Class", "Genotype"). These variables should represent the groups that are going to be compared. It will then create new columns containing the descriptive stats that were chosen or default to all of them (e.g. n, median, q1, q3, mean, SD, and SEM) for each numerical (cell) variables that was provided (e.g., "Brain_T_Cell_Counts", "Brain_Microglia_Counts","Brain_Neutrophil_Counts"). 

```{r}
descriptive_stats_example <- descriptive_data(
  data = combined_data2, # data frame
  group_variables = c("Sex","ZT", "Age_Class", "Genotype"),  #categorical group variables
  cell_variables = c("Brain_T_Cell_Counts", "Brain_Microglia_Counts","Brain_Neutrophil_Counts"), # numerical cell variables
  stats= c("n", "median", "q1", "q3", "mean", "sd", "sem")) # descriptive stats

knitr::kable(descriptive_stats_example)

```
