---
title: "Make_model_table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Make_model_table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bim)
```

## PRELIMINARIES 
```{r}
library(readr)
library(dplyr)
library(skimr)
```

## DATA WRANGLING 
```{r}
data("UT27_Meninges_Output_Data")
meninges <- UT27_Meninges_Output_Data

data("UT27_Brain_Output_Data")
brain <- UT27_Brain_Output_Data

brain_select <- brain %>%
  select(Date, `Mouse #`,`Cage #`, ZT, Sex, Group, `Microglia Counts`, `Neutrophil Counts`, `T-Cell Counts`, `CD8+ T Cell Counts`, `CD4+ T Cell Counts`, `B Cell Counts`)

meninges_select <- meninges %>%
  select(Date, `Mouse #`,`Cage #`, ZT, Sex, Group, `Neutrophil Cells`, `T Cell Cells`, `CD8 Cells`, `CD4 Cells`, `B Cell Cells`)

meninges_select <- meninges_select %>%
  rename(
    `Meninges Neutrophil Counts` = `Neutrophil Cells`,
    `Meninges T-Cell Counts` = `T Cell Cells`,
    `Meninges CD8+ T Cell Counts` = `CD8 Cells`,
    `Meninges CD4+ T Cell Counts` = `CD4 Cells`,
    `Meninges B Cell Counts` = `B Cell Cells`
  )

brain_select <- brain_select %>%
  rename(
    `Brain Microglia Counts` = `Microglia Counts`,
    `Brain Neutrophil Counts` = `Neutrophil Counts`,
    `Brain T-Cell Counts` = `T-Cell Counts`,
    `Brain CD8+ T Cell Counts` = `CD8+ T Cell Counts`,
    `Brain CD4+ T Cell Counts` = `CD4+ T Cell Counts`,
    `Brain B Cell Counts` = `B Cell Counts`
  )

brain_select$`Mouse #` <- paste(brain_select$`Cage #`, brain_select$`Mouse #`, sep = "-")
meninges_select$`Mouse #` <- paste(meninges_select$`Cage #`, meninges_select$`Mouse #`, sep = "-")

combined_data <- brain_select %>%
  full_join(meninges_select) 

combined_data <- na.omit(combined_data)
combined_data <- combined_data %>%
  filter(!(Group %in% c("2-WT-9", "3-3xTg-9", "5-3xTg-18")))
```

# COEFFICIENT HEATMAP
This will generate a heatmap that gives an overview of which cell types correlate with which cell types. Ones that positively correlate will be represented by a blue box, negatively correlate by a red box, and don't significantly correlate by a grey box. In this example, we can see that all T cell and CD8+ T cell counts correlate with each other, even between tissues. This should inform which are the best predictors to use in a model.

This heatmap validates our variable choices by highlighting meaningful correlations between cell types, helping us avoid redundancy and multicollinearity in downstream modeling. It ensures our analysis focuses on biologically relevant relationships worth exploring further.

```{r}
library(corrplot)
numeric_data <- combined_data[, sapply(combined_data, is.numeric)]
corrplot::corrplot(cor(numeric_data, use = "complete.obs"), method = "color")
```

## MAKE MODEL TABLE EXAMPLE
This function takes an input of a linear model and generates a table showing the Estimates, SE, p value, and F score for each variable used in that model, allowing you to see how good of a predictor each variable is. This was done below for both the full and reduced model from the choose_best_model function.

```{r}
library(car)

#evaluate models
brainCD8_full_table <- make_model_table(brain_CD8_model)
brainCD8_reduced_table <- make_model_table(brain_CD8_model_reduced)

print(brainCD8_full_table)
print(brainCD8_reduced_table)

```
