---
title: "Plot_model_with_ci"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plot_model_with_ci}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bim)
```

## PRELIMINARIES 

```{r}
library(readr)
library(dplyr)
library(skimr)
```

## DATA WRANGLING 

```{r}
data("UT27_Meninges_Output_Data")
meninges <- UT27_Meninges_Output_Data

data("UT27_Brain_Output_Data")
brain <- UT27_Brain_Output_Data

brain_select <- brain %>%
  select(Date, `Mouse #`,`Cage #`, ZT, Sex, Group, `Microglia Counts`, `Neutrophil Counts`, `T-Cell Counts`, `CD8+ T Cell Counts`, `CD4+ T Cell Counts`, `B Cell Counts`)

meninges_select <- meninges %>%
  select(Date, `Mouse #`,`Cage #`, ZT, Sex, Group, `Neutrophil Cells`, `T Cell Cells`, `CD8 Cells`, `CD4 Cells`, `B Cell Cells`)

meninges_select <- meninges_select %>%
  rename(
    `Meninges Neutrophil Counts` = `Neutrophil Cells`,
    `Meninges T-Cell Counts` = `T Cell Cells`,
    `Meninges CD8+ T Cell Counts` = `CD8 Cells`,
    `Meninges CD4+ T Cell Counts` = `CD4 Cells`,
    `Meninges B Cell Counts` = `B Cell Cells`
  )

brain_select <- brain_select %>%
  rename(
    `Brain Microglia Counts` = `Microglia Counts`,
    `Brain Neutrophil Counts` = `Neutrophil Counts`,
    `Brain T-Cell Counts` = `T-Cell Counts`,
    `Brain CD8+ T Cell Counts` = `CD8+ T Cell Counts`,
    `Brain CD4+ T Cell Counts` = `CD4+ T Cell Counts`,
    `Brain B Cell Counts` = `B Cell Counts`
  )

brain_select$`Mouse #` <- paste(brain_select$`Cage #`, brain_select$`Mouse #`, sep = "-")
meninges_select$`Mouse #` <- paste(meninges_select$`Cage #`, meninges_select$`Mouse #`, sep = "-")

combined_data <- brain_select %>%
  full_join(meninges_select) 

combined_data <- na.omit(combined_data)
combined_data <- combined_data %>%
  filter(!(Group %in% c("2-WT-9", "3-3xTg-9", "5-3xTg-18")))
```

# COEFFICIENT HEATMAP
DESCRIPTION: This will generate a heatmap that gives an overview of which cell types correlate with which cell types. Ones that positively correlate will be represented by a blue box, negatively correlate by a red box, and don't significantly correlate by a grey box. In this example, we can see that all T cell and CD8+ T cell counts correlate with each other, even between tissues. This should inform which are the best predictors to use in a model.

```{r}
library(corrplot)
numeric_data <- combined_data[, sapply(combined_data, is.numeric)]
corrplot::corrplot(cor(numeric_data, use = "complete.obs"), method = "color")
```

## PLOT MODEL WITH CI EXAMPLE

Ex. Use this:
DESCRIPTION: This function takes several inputs: model (the linear model used for this data), data (the dataframe that model is based off of), y_var (the variable that you are modeling), x_var (the predictive variable of interest), and group_var (The column in the dataframe that you want to compare between groups). This allows us to plot twoi cell types against each other and compare how they interact in different groups.

```{r}
library(ggplot2)

bCD8xmCD8_plot <- plot_model_with_ci(
  model = brain_CD8_model_reduced,
  data = combined_data,
  y_var = "Brain CD8+ T Cell Counts",
  x_var = "Meninges CD8+ T Cell Counts",
  group_var = "Group"
)

print(bCD8xmCD8_plot)
```
